---
title: "FCE Portfolio Risk & Monte Carlo Simulation"
author: "Brian Connolly | MS Quantitative Economics, UConn 2024"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    theme:
      version: 4
      bootswatch: flatly
      primary: "#2D6A4F"
      secondary: "#52B788"
    css: styles.css
    navbar:
      - title: "CECL Dashboard"
        href: "https://brianconnolly-models-portfolio.github.io/fce-risk-dashboard/FCE_Dashboard.html"
        target: "_blank"
        align: right
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE,
                      message = FALSE, fig.retina = 2)

library(flexdashboard)
library(tidyverse)
library(scales)
library(plotly)
library(DT)

# ── FCE color palette ──────────────────────────────────────
fce_green  <- "#2D6A4F"
fce_mid    <- "#52B788"
fce_light  <- "#95D5B2"
fce_orange <- "#F4A261"
fce_red    <- "#C1121F"
fce_dark   <- "#1B4332"
fce_cream  <- "#F8F4E8"

# ── Shared ggplot theme ────────────────────────────────────
theme_fce <- function(base = 12) {
  theme_minimal(base_size = base) +
    theme(
      plot.title       = element_text(face = "bold",
                                      size = base + 2,
                                      color = fce_dark),
      plot.subtitle    = element_text(size = base - 1,
                                      color = "gray40"),
      plot.caption     = element_text(size = base - 3,
                                      color = "gray55"),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_line(color = "gray92"),
      axis.title       = element_text(size = base - 1,
                                      color = "gray35"),
      plot.background  = element_rect(fill = "white",
                                      color = NA),
      plot.margin      = margin(12, 12, 12, 12)
    )
}

# ── Key metrics (pre-computed for KPI cards) ───────────────
el_mean      <- mean(sim_results$total_loss)
el_median    <- median(sim_results$total_loss)
var_75       <- quantile(sim_results$total_loss, 0.75)
var_95       <- quantile(sim_results$total_loss, 0.95)
var_99       <- quantile(sim_results$total_loss, 0.99)
cvar_95      <- mean(sim_results$total_loss[
                  sim_results$total_loss >= var_95])
cecl_reserve <- 8100000
prob_ok      <- mean(sim_results$total_loss <= cecl_reserve)
buffer       <- cecl_reserve - var_99
ag_factor    <- round(1.48 / 21.4, 4)
```
```{css, echo=FALSE}
/* ── Typography ─────────────────────────────────────────── */
@import url('https://fonts.googleapis.com/css2?family=DM+Serif+Display&family=DM+Sans:wght@300;400;500;600&family=DM+Mono&display=swap');

body, .dashboard-page {
  font-family: 'DM Sans', sans-serif;
  background-color: #F8F4E8;
}

/* ── Navbar ──────────────────────────────────────────────── */
.navbar {
  background-color: #1B4332 !important;
  border-bottom: 3px solid #2D6A4F;
}
.navbar-brand {
  font-family: 'DM Serif Display', serif !important;
  font-size: 1.15rem !important;
  letter-spacing: 0.03em;
  color: #F8F4E8 !important;
}
.navbar-nav > li > a {
  color: #95D5B2 !important;
  font-size: 0.82rem;
  font-weight: 500;
  letter-spacing: 0.04em;
  text-transform: uppercase;
}
.navbar-nav > li > a:hover {
  color: #F8F4E8 !important;
}

/* ── Tab navigation ─────────────────────────────────────── */
.nav-tabs-custom > .nav-tabs > li > a {
  font-family: 'DM Sans', sans-serif;
  font-size: 0.8rem;
  font-weight: 600;
  letter-spacing: 0.06em;
  text-transform: uppercase;
  color: #2D6A4F;
}
.nav-tabs-custom > .nav-tabs > li.active > a {
  color: #1B4332;
  border-top: 3px solid #2D6A4F;
}

/* ── Value boxes ─────────────────────────────────────────── */
.value-box {
  border-radius: 6px;
  border: none;
  box-shadow: 0 2px 8px rgba(27,67,50,0.10);
}
.value-box .value-box-value {
  font-family: 'DM Serif Display', serif;
  font-size: 1.7rem;
}
.value-box .value-box-caption {
  font-size: 0.72rem;
  font-weight: 600;
  letter-spacing: 0.07em;
  text-transform: uppercase;
  opacity: 0.85;
}

/* ── Chart panels ───────────────────────────────────────── */
.chart-wrapper, .chart-stage {
  background: #ffffff;
  border-radius: 6px;
  box-shadow: 0 2px 8px rgba(27,67,50,0.07);
  padding: 4px;
}

/* ── Section headers ────────────────────────────────────── */
h3.section-header {
  font-family: 'DM Serif Display', serif;
  color: #1B4332;
  border-bottom: 2px solid #95D5B2;
  padding-bottom: 6px;
  margin-bottom: 16px;
  font-size: 1.1rem;
}

/* ── Methodology boxes ──────────────────────────────────── */
.method-box {
  background: white;
  border-left: 4px solid #2D6A4F;
  border-radius: 0 6px 6px 0;
  padding: 16px 20px;
  margin-bottom: 14px;
  box-shadow: 0 1px 4px rgba(27,67,50,0.08);
}
.method-box h4 {
  font-family: 'DM Serif Display', serif;
  color: #1B4332;
  margin-top: 0;
  margin-bottom: 8px;
  font-size: 1rem;
}
.method-box p, .method-box li {
  font-size: 0.88rem;
  color: #333;
  line-height: 1.6;
  margin-bottom: 4px;
}
.method-box code {
  font-family: 'DM Mono', monospace;
  background: #f0f7f4;
  color: #2D6A4F;
  padding: 1px 5px;
  border-radius: 3px;
  font-size: 0.82rem;
}

/* ── Stat badges ────────────────────────────────────────── */
.stat-badge {
  display: inline-block;
  background: #f0f7f4;
  color: #1B4332;
  border: 1px solid #95D5B2;
  border-radius: 4px;
  padding: 3px 10px;
  font-family: 'DM Mono', monospace;
  font-size: 0.82rem;
  font-weight: 600;
  margin: 2px;
}

/* ── DT tables ──────────────────────────────────────────── */
.dataTables_wrapper {
  font-size: 0.84rem;
}
table.dataTable thead th {
  background-color: #1B4332;
  color: #F8F4E8;
  font-family: 'DM Sans', sans-serif;
  font-size: 0.75rem;
  letter-spacing: 0.05em;
  text-transform: uppercase;
}
```

# Methodology {data-icon="fa-flask"}

## Row {data-height=120}

### {.no-padding}

<div style="background: linear-gradient(135deg, #1B4332 0%, #2D6A4F 60%, #52B788 100%);
            padding: 28px 36px; border-radius: 6px; color: white; margin-bottom: 4px;">
  <div style="font-family: 'DM Serif Display', serif; font-size: 1.8rem;
              letter-spacing: 0.02em; margin-bottom: 6px;">
    FCE Portfolio Risk &amp; Monte Carlo Simulation
  </div>
  <div style="font-size: 0.88rem; opacity: 0.85; font-weight: 300; letter-spacing: 0.04em;">
    A full-stack credit risk model — logistic regression PD estimation,
    agricultural domain calibration, and macro-driven loss simulation
  </div>
  <div style="margin-top: 12px;">
    <span class="stat-badge" style="background:rgba(255,255,255,0.15);
          color:white; border-color:rgba(255,255,255,0.3);">
      2.26M Training Loans
    </span>
    <span class="stat-badge" style="background:rgba(255,255,255,0.15);
          color:white; border-color:rgba(255,255,255,0.3);">
      AUC 0.70
    </span>
    <span class="stat-badge" style="background:rgba(255,255,255,0.15);
          color:white; border-color:rgba(255,255,255,0.3);">
      10,000 Scenarios
    </span>
    <span class="stat-badge" style="background:rgba(255,255,255,0.15);
          color:white; border-color:rgba(255,255,255,0.3);">
      FRED 2000–2025
    </span>
    <span class="stat-badge" style="background:rgba(255,255,255,0.15);
          color:white; border-color:rgba(255,255,255,0.3);">
      USDA ERS Validated
    </span>
  </div>
</div>

## Row {data-height=480}

### Step 1 — Data Acquisition {data-width=250}

<div class="method-box">
<h4>📦 Lending Club Loan Data</h4>
<p>2,260,668 US consumer loans with observed default outcomes sourced from Kaggle. Exceptional data quality — less than 0.1% missing values across key predictors.</p>
<ul>
<li>Binary target: Charged Off / Late 31-120 days = Default</li>
<li>Excluded ambiguous statuses (Current, Grace Period)</li>
<li>Final training set: <strong>1,327,062 loans</strong></li>
<li>Observed default rate: <strong>21.4%</strong></li>
</ul>
</div>

<div class="method-box">
<h4>📈 FRED Macro Series</h4>
<p>Five Federal Reserve economic series pulled via <code>fredr</code> package, covering 2000–2025.</p>
<ul>
<li>Fed Funds Rate — debt service driver</li>
<li>Corn Prices (WPU012) — Row Crop revenue</li>
<li>Milk Prices (WPU0231) — Dairy revenue</li>
<li>Farmland Values — collateral proxy</li>
<li>Ag Delinquency Rate (DRALACBS) — <strong>1.48%</strong> current</li>
</ul>
</div>

<div class="method-box">
<h4>🌾 USDA ERS Farm Income</h4>
<p>Net Farm Income data 2000–2026 used for macro context and stress scenario calibration.</p>
<ul>
<li>2022 peak: <strong>$182B</strong> (all-time high)</li>
<li>2015–2016 crash: <strong>–32%</strong> stress analog</li>
<li>2026 estimate: <strong>$153B</strong> (normalization)</li>
</ul>
</div>

### Step 2 — Logistic Regression PD Model {data-width=250}

<div class="method-box">
<h4>🧮 Model Specification</h4>
<p>Logistic regression trained on 80% of definitive-outcome loans, validated on held-out 20% test set.</p>
<ul>
<li>13 predictors — all known at origination</li>
<li>Family: <code>binomial(link = "logit")</code></li>
<li>Train set: <strong>1,061,651 loans</strong></li>
<li>Test set: <strong>265,411 loans</strong></li>
<li>Test AUC: <strong>0.7012</strong></li>
</ul>
</div>

<div class="method-box">
<h4>📊 Key Predictors</h4>
<p>All 13 predictors statistically significant (p &lt; 0.05). Top drivers by odds ratio:</p>
<ul>
<li><code>small_biz</code> — 1.43× default odds</li>
<li><code>log_loan_amnt</code> — 1.22× per log unit</li>
<li><code>home_risk</code> — 1.20× (renter vs owner)</li>
<li><code>grade_numeric</code> — 1.18× per grade step</li>
<li><code>log_annual_inc</code> — 0.75× (protective)</li>
</ul>
</div>

<div class="method-box">
<h4>🎯 PD Staircase Validation</h4>
<p>Predicted PDs follow a perfect monotonic staircase across credit grades — confirming model discriminatory power.</p>
<ul>
<li>Grade A: <strong>9.2%</strong> predicted vs 6.5% actual</li>
<li>Grade D: <strong>30.9%</strong> predicted vs 32.1% actual</li>
<li>Grade G: <strong>61.6%</strong> predicted vs 50.2% actual</li>
<li>Northeast premium: <strong>+0.8%</strong> vs national</li>
</ul>
</div>

### Step 3 — Domain Calibration & Monte Carlo {data-width=250}

<div class="method-box">
<h4>⚖️ Agricultural Scaling Factor</h4>
<p>Logistic regression PDs reflect consumer lending default rates. A domain calibration factor translates to agricultural lending reality.</p>
<ul>
<li>FRED Ag Delinquency: <strong>1.48%</strong></li>
<li>LC Observed Default: <strong>21.4%</strong></li>
<li>Factor: <strong>1.48 ÷ 21.4 = 0.069</strong></li>
<li>Both inputs from Step 1 data pulls</li>
<li>Conservative — delinquency slightly overstates default</li>
</ul>
</div>

<div class="method-box">
<h4>🎲 Monte Carlo Engine</h4>
<p>10,000 independent economic scenarios, each drawing from 25 years of observed FRED history.</p>
<ul>
<li>Bootstrap: sample with replacement from FRED</li>
<li>Macro multipliers per segment sensitivity</li>
<li>Bernoulli draw: <code>rbinom(n(), 1, stressed_pd)</code></li>
<li>Loss: <code>EAD × LGD × default indicator</code></li>
</ul>
</div>

<div class="method-box">
<h4>📐 Stress Multiplier Logic</h4>
<p>Each macro scenario adjusts base PD through segment-specific sensitivities:</p>
<ul>
<li>Rate stress: +8% PD per 1% above 3.72%</li>
<li>Corn stress: Row Crop sensitivity 1.80×</li>
<li>Milk stress: Dairy sensitivity 1.80×</li>
<li>PD capped at 95% — no loan is certain default</li>
</ul>
</div>

### Step 4 — Outputs & Validation {data-width=250}

<div class="method-box">
<h4>✅ Convergence Validation</h4>
<p>Cumulative mean loss stabilizes well before simulation 5,000 — confirming 10,000 simulations produces stable estimates.</p>
<ul>
<li>Mean converges to: <strong>`r dollar(el_mean)`</strong></li>
<li>Median–mean gap: <strong>`r dollar(abs(el_mean - el_median))`</strong></li>
<li>Symmetric distribution — no extreme skew</li>
</ul>
</div>

<div class="method-box">
<h4>🔗 Model Chain Integrity</h4>
<p>Every number in this dashboard traces back to a real data source — no assumptions without derivation.</p>
<ul>
<li>PDs → Lending Club 2.26M loan outcomes</li>
<li>Scaling factor → FRED DRALACBS series</li>
<li>Macro scenarios → FRED bootstrapped history</li>
<li>Stress analogs → USDA ERS farm income 2015–16</li>
</ul>
</div>

<div class="method-box">
<h4>📋 Limitations & Extensions</h4>
<p>Documented limitations for regulatory transparency:</p>
<ul>
<li>PD model trained on consumer, not ag loans</li>
<li>LGD estimates are segment assumptions, not observed</li>
<li>Correlation across loans not explicitly modeled</li>
<li><strong>Next step:</strong> Re-train on FCE loan-level data</li>
<li><strong>Next step:</strong> Add Gaussian copula for correlation</li>
</ul>
</div>

# Loss Distribution & VaR {data-icon="fa-chart-area"}

## Row {data-height=100}

### Expected Loss {.value-box}
```{r}
valueBox(
  value   = dollar(el_mean, accuracy = 1),
  caption = "Expected Loss — Mean of 10,000 Scenarios",
  icon    = "fa-dollar-sign",
  color   = fce_green
)
```

### 95th Pct VaR {.value-box}
```{r}
valueBox(
  value   = dollar(var_95, accuracy = 1),
  caption = "Value at Risk — 95th Percentile",
  icon    = "fa-exclamation-triangle",
  color   = fce_orange
)
```

### 99th Pct Loss {.value-box}
```{r}
valueBox(
  value   = dollar(var_99, accuracy = 1),
  caption = "Severe Stress — 99th Percentile",
  icon    = "fa-fire",
  color   = fce_red
)
```

### Reserve Buffer {.value-box}
```{r}
valueBox(
  value   = dollar(buffer, accuracy = 1),
  caption = "Reserve Buffer Above 99th Pct Loss",
  icon    = "fa-shield-alt",
  color   = ifelse(buffer > 0, fce_green, fce_red)
)
```

## Row {data-height=420}

### Loss Distribution — 10,000 Monte Carlo Scenarios {data-width=650}
```{r fig.height=5}
var_95_cutoff <- quantile(sim_results$total_loss / 1e6, 0.95)

p1 <- ggplot(sim_results, aes(x = total_loss / 1e6)) +
  geom_histogram(
    aes(fill = after_stat(x) > var_95_cutoff),
    bins = 70, color = "white", linewidth = 0.15
  ) +
  scale_fill_manual(
    values = c("FALSE" = fce_light, "TRUE" = fce_red),
    guide  = "none"
  ) +
  geom_vline(xintercept = cecl_reserve / 1e6,
             color = fce_dark, linetype = "dashed",
             linewidth = 1.1) +
  geom_vline(xintercept = el_mean / 1e6,
             color = fce_green, linetype = "solid",
             linewidth = 1.1) +
  geom_vline(xintercept = var_95 / 1e6,
             color = fce_red, linetype = "solid",
             linewidth = 1.1) +
  annotate("text",
           x = cecl_reserve / 1e6, y = Inf,
           label = paste0("Reserve\n$",
                          round(cecl_reserve/1e6, 1), "M"),
           vjust = 1.3, hjust = 1.1,
           color = fce_dark, size = 3.2,
           fontface = "bold") +
  annotate("text",
           x = el_mean / 1e6, y = Inf,
           label = paste0("EL\n$",
                          round(el_mean/1e6, 2), "M"),
           vjust = 1.3, hjust = -0.1,
           color = fce_green, size = 3.2,
           fontface = "bold") +
  annotate("text",
           x = var_95 / 1e6, y = Inf,
           label = paste0("VaR 95\n$",
                          round(var_95/1e6, 1), "M"),
           vjust = 1.3, hjust = -0.1,
           color = fce_red, size = 3.2,
           fontface = "bold") +
  scale_x_continuous(
    labels = dollar_format(suffix = "M", accuracy = 0.1)
  ) +
  scale_y_continuous(labels = comma_format()) +
  labs(
    title    = "Portfolio Loss Distribution",
    subtitle = paste0("Red = worst 5% of scenarios | ",
                      "Current reserve covers ",
                      percent(prob_ok, accuracy = 0.1),
                      " of simulations"),
    x = "Total Portfolio Loss ($M)",
    y = "Number of Scenarios",
    caption = "Source: FRED bootstrapped macro inputs | Logistic Regression PD | 0.069 ag calibration factor"
  ) +
  theme_fce()

ggplotly(p1, tooltip = c("x", "y")) %>%
  layout(hoverlabel = list(bgcolor = "white",
                           font = list(family = "DM Sans")))
```

### Reserve Adequacy by Confidence Level {data-width=350}
```{r fig.height=5}
reserve_levels <- tibble(
  percentile = c(50, 60, 70, 75, 80, 85, 90, 95, 99),
  loss       = quantile(sim_results$total_loss,
                        c(.50,.60,.70,.75,.80,
                          .85,.90,.95,.99))
) %>%
  mutate(
    label    = paste0(percentile, "th"),
    adequate = loss <= cecl_reserve
  )

p2 <- ggplot(reserve_levels,
             aes(x    = reorder(label, percentile),
                 y    = loss / 1e6,
                 fill = adequate,
                 text = paste0(percentile, "th pct: ",
                               dollar(loss)))) +
  geom_col(width = 0.7) +
  geom_hline(yintercept = cecl_reserve / 1e6,
             color = fce_dark, linetype = "dashed",
             linewidth = 1.1) +
  geom_text(aes(label = dollar(loss/1e6,
                               accuracy = 0.1,
                               suffix = "M")),
            vjust = -0.4, size = 2.8,
            fontface = "bold") +
  scale_fill_manual(
    values = c("TRUE" = fce_green, "FALSE" = fce_red),
    labels = c("TRUE" = "Adequate", "FALSE" = "Insufficient"),
    name   = NULL
  ) +
  scale_y_continuous(
    labels = dollar_format(suffix = "M", accuracy = 1),
    expand = expansion(mult = c(0, 0.18))
  ) +
  labs(
    title    = "Required Capital by Confidence Level",
    subtitle = paste0("CECL reserve: $",
                      round(cecl_reserve/1e6, 1), "M"),
    x = "Loss Percentile", y = "Required Reserve ($M)",
    caption = "Source: FCE Monte Carlo Simulation"
  ) +
  theme_fce() +
  theme(legend.position = "top")

ggplotly(p2, tooltip = "text") %>%
  layout(hoverlabel = list(bgcolor = "white"))
```

## Row {data-height=200}

### Convergence Check — 10,000 Simulations {data-width=650}
```{r fig.height=2.8}
convergence <- sim_results %>%
  mutate(cumulative_mean = cummean(total_loss))

p_conv <- ggplot(convergence,
                 aes(x = sim_id,
                     y = cumulative_mean / 1e6)) +
  geom_line(color = fce_green, linewidth = 0.8) +
  geom_hline(yintercept = el_mean / 1e6,
             color = fce_red, linetype = "dashed",
             linewidth = 0.7) +
  scale_x_continuous(labels = comma_format()) +
  scale_y_continuous(
    labels = dollar_format(suffix = "M", accuracy = 0.1)
  ) +
  labs(
    title    = "Monte Carlo Convergence",
    subtitle = "Estimate stabilizes — 10,000 simulations confirmed sufficient",
    x = "Simulations Run", y = "Cumulative Mean Loss",
    caption  = "Dashed = final expected loss estimate"
  ) +
  theme_fce()

ggplotly(p_conv, tooltip = c("x", "y")) %>%
  layout(hoverlabel = list(bgcolor = "white"))
```

### Loss Distribution Summary {data-width=350}
```{r}
tibble(
  Metric = c("Expected Loss (Mean)",
             "Median Loss (50th)",
             "Elevated (75th)",
             "Value at Risk (95th)",
             "Severe Stress (99th)",
             "Exp. Shortfall (CVaR)",
             "Current CECL Reserve",
             "Reserve Buffer (99th)"),
  Value  = c(dollar(el_mean),
             dollar(el_median),
             dollar(var_75),
             dollar(var_95),
             dollar(var_99),
             dollar(cvar_95),
             dollar(cecl_reserve),
             dollar(buffer))
) %>%
  datatable(
    rownames  = FALSE,
    options   = list(dom = "t", pageLength = 10,
                     scrollY = "160px"),
    class     = "compact stripe"
  ) %>%
  formatStyle(
    "Metric",
    fontWeight = "600",
    fontSize   = "0.83rem"
  ) %>%
  formatStyle(
    "Value",
    fontFamily = "DM Mono, monospace",
    fontSize   = "0.83rem",
    color      = fce_dark
  )
```

# Segment Risk {data-icon="fa-layer-group"}

## Row {data-height=100}

### Dairy Avg Loss {.value-box}
```{r}
valueBox(
  value   = dollar(mean(sim_results$dairy_loss)),
  caption = "Dairy — Avg Loss (39% of Total)",
  icon    = "fa-tint",
  color   = fce_green
)
```

### Livestock Avg Loss {.value-box}
```{r}
valueBox(
  value   = dollar(mean(sim_results$livestock_loss)),
  caption = "Livestock — Avg Loss (20% of Total)",
  icon    = "fa-horse",
  color   = fce_mid
)
```

### Row Crop Avg Loss {.value-box}
```{r}
valueBox(
  value   = dollar(mean(sim_results$row_crop_loss)),
  caption = "Row Crop — Avg Loss (18% of Total)",
  icon    = "fa-seedling",
  color   = fce_orange
)
```

### Agri-Business Avg Loss {.value-box}
```{r}
valueBox(
  value   = dollar(mean(sim_results$agribiz_loss)),
  caption = "Agri-Business — Avg Loss (11% of Total)",
  icon    = "fa-industry",
  color   = "#6B7280"
)
```

## Row {data-height=450}

### Segment Loss Distribution {data-width=550}
```{r fig.height=5}
seg_long <- sim_results %>%
  select(sim_id,
         Dairy           = dairy_loss,
         Livestock       = livestock_loss,
         `Row Crop`      = row_crop_loss,
         `Agri-Business` = agribiz_loss) %>%
  pivot_longer(-sim_id,
               names_to  = "segment",
               values_to = "loss")

p3 <- ggplot(
  seg_long,
  aes(x    = reorder(segment, loss, median),
      y    = loss / 1e6,
      fill = segment,
      text = paste0(segment, "\nLoss: ",
                    dollar(loss, accuracy = 1)))
) +
  geom_boxplot(
    outlier.shape = 16,
    outlier.size  = 0.4,
    outlier.alpha = 0.25,
    width         = 0.55
  ) +
  coord_flip() +
  scale_fill_manual(
    values = c(
      "Dairy"         = fce_green,
      "Row Crop"      = fce_orange,
      "Livestock"     = fce_mid,
      "Agri-Business" = "#D1D5DB"
    ),
    guide = "none"
  ) +
  scale_y_continuous(
    labels = dollar_format(suffix = "M", accuracy = 0.1)
  ) +
  labs(
    title    = "Segment Loss Distribution",
    subtitle = "Box = 25th–75th pct | Line = median | Whiskers = 5th–95th pct",
    x = NULL, y = "Segment Loss ($M)",
    caption  = "Source: FCE Monte Carlo | 10,000 scenarios"
  ) +
  theme_fce()

ggplotly(p3, tooltip = "text") %>%
  layout(hoverlabel = list(bgcolor = "white"))
```

### Segment Risk Summary Table {data-width=450}
```{r}
seg_summary <- tibble(
  Segment = c("Dairy", "Livestock",
              "Row Crop", "Agri-Business"),
  Portfolio_Wt = c("28%", "18%", "15%", "14%"),
  Avg_Loss = c(
    mean(sim_results$dairy_loss),
    mean(sim_results$livestock_loss),
    mean(sim_results$row_crop_loss),
    mean(sim_results$agribiz_loss)
  ),
  VaR_95 = c(
    quantile(sim_results$dairy_loss,    0.95),
    quantile(sim_results$livestock_loss, 0.95),
    quantile(sim_results$row_crop_loss,  0.95),
    quantile(sim_results$agribiz_loss,   0.95)
  ),
  Share_of_EL = c(
    mean(sim_results$dairy_loss)    / el_mean,
    mean(sim_results$livestock_loss) / el_mean,
    mean(sim_results$row_crop_loss)  / el_mean,
    mean(sim_results$agribiz_loss)   / el_mean
  )
) %>%
  mutate(
    Conc_Ratio = parse_number(Portfolio_Wt) / 100,
    Conc_Ratio = round(Share_of_EL / Conc_Ratio, 2),
    Avg_Loss   = dollar(Avg_Loss),
    VaR_95     = dollar(VaR_95),
    Share_of_EL = percent(Share_of_EL, accuracy = 0.1)
  ) %>%
  select(Segment, Portfolio_Wt,
         `Avg Loss`    = Avg_Loss,
         `VaR 95th`    = VaR_95,
         `EL Share`    = Share_of_EL,
         `Conc. Ratio` = Conc_Ratio)

datatable(
  seg_summary,
  rownames = FALSE,
  options  = list(dom = "t", pageLength = 6),
  class    = "compact stripe"
) %>%
  formatStyle(
    "Conc. Ratio",
    backgroundColor = styleInterval(
      c(1.0, 1.3),
      c("#f0f7f4", "#fff3cd", "#fde8e8")
    ),
    fontWeight = "bold"
  ) %>%
  formatStyle(
    columns    = names(seg_summary),
    fontSize   = "0.84rem"
  )
```

### Key Insight {data-width=0}
```{r}
# Concentration ratio note rendered inline above table
```

# Macro Sensitivity {data-icon="fa-chart-line"}

## Row {data-height=100}

### Rate Environment {.value-box}
```{r}
valueBox(
  value   = "3.72%",
  caption = "Current Fed Funds Rate — Baseline",
  icon    = "fa-university",
  color   = fce_green
)
```

### Corn Price Index {.value-box}
```{r}
valueBox(
  value   = "165.6",
  caption = "Corn Price Index (WPU012) — Current",
  icon    = "fa-seedling",
  color   = fce_mid
)
```

### Milk Price Index {.value-box}
```{r}
valueBox(
  value   = "286.9",
  caption = "Milk Price Index (WPU0231) — Favorable",
  icon    = "fa-tint",
  color   = fce_orange
)
```

### Ag Delinquency {.value-box}
```{r}
valueBox(
  value   = "1.48%",
  caption = "FRED Ag Delinquency Rate — DRALACBS",
  icon    = "fa-percentage",
  color   = fce_dark
)
```

## Row {data-height=440}

### Interest Rate vs Portfolio Loss {data-width=400}
```{r fig.height=4.8}
p_rate <- ggplot(
  sim_results,
  aes(x    = rate_scenario,
      y    = total_loss / 1e6,
      text = paste0("Rate: ",
                    round(rate_scenario, 2), "%\n",
                    "Loss: ",
                    dollar(total_loss, accuracy = 1)))
) +
  geom_point(alpha = 0.04, color = fce_green,
             size = 0.7) +
  geom_smooth(method = "lm", color = fce_red,
              se = TRUE, linewidth = 1) +
  geom_vline(xintercept = 3.72,
             color = fce_dark, linetype = "dashed",
             linewidth = 0.8) +
  annotate("text", x = 3.72, y = Inf,
           label = "Current\n3.72%",
           vjust = 1.4, hjust = -0.1,
           size = 3, color = fce_dark,
           fontface = "bold") +
  scale_x_continuous(
    labels = function(x) paste0(x, "%")
  ) +
  scale_y_continuous(
    labels = dollar_format(suffix = "M", accuracy = 1)
  ) +
  labs(
    title    = "Interest Rate vs Portfolio Loss",
    subtitle = "Higher rates amplify PD via debt service stress",
    x = "Fed Funds Rate (%)",
    y = "Total Loss ($M)",
    caption  = "Red line = OLS trend | Shading = 95% confidence interval"
  ) +
  theme_fce()

ggplotly(p_rate, tooltip = "text") %>%
  layout(hoverlabel = list(bgcolor = "white"))
```

### Corn Price vs Portfolio Loss {data-width=330}
```{r fig.height=4.8}
p_corn <- ggplot(
  sim_results,
  aes(x    = corn_scenario * 100,
      y    = total_loss / 1e6,
      text = paste0("Corn chg: ",
                    round(corn_scenario * 100, 1), "%\n",
                    "Loss: ",
                    dollar(total_loss, accuracy = 1)))
) +
  geom_point(alpha = 0.04, color = fce_orange,
             size = 0.7) +
  geom_smooth(method = "lm", color = fce_red,
              se = TRUE, linewidth = 1) +
  geom_vline(xintercept = 0,
             color = "gray50", linetype = "dashed",
             linewidth = 0.7) +
  scale_x_continuous(
    labels = function(x) paste0(x, "%")
  ) +
  scale_y_continuous(
    labels = dollar_format(suffix = "M", accuracy = 1)
  ) +
  labs(
    title    = "Corn Price Change vs Loss",
    subtitle = "Row Crop sensitivity: 1.80×",
    x = "Monthly Corn Price Change (%)",
    y = "Total Loss ($M)",
    caption  = "Negative = price decline = higher Row Crop PD"
  ) +
  theme_fce()

ggplotly(p_corn, tooltip = "text") %>%
  layout(hoverlabel = list(bgcolor = "white"))
```

### Milk Price vs Portfolio Loss {data-width=330}
```{r fig.height=4.8}
p_milk <- ggplot(
  sim_results,
  aes(x    = milk_scenario * 100,
      y    = total_loss / 1e6,
      text = paste0("Milk chg: ",
                    round(milk_scenario * 100, 1), "%\n",
                    "Loss: ",
                    dollar(total_loss, accuracy = 1)))
) +
  geom_point(alpha = 0.04, color = "#6B4C9A",
             size = 0.7) +
  geom_smooth(method = "lm", color = fce_red,
              se = TRUE, linewidth = 1) +
  geom_vline(xintercept = 0,
             color = "gray50", linetype = "dashed",
             linewidth = 0.7) +
  scale_x_continuous(
    labels = function(x) paste0(x, "%")
  ) +
  scale_y_continuous(
    labels = dollar_format(suffix = "M", accuracy = 1)
  ) +
  labs(
    title    = "Milk Price Change vs Loss",
    subtitle = "Dairy sensitivity: 1.80× | Largest absolute segment",
    x = "Monthly Milk Price Change (%)",
    y = "Total Loss ($M)",
    caption  = "Negative = price decline = higher Dairy PD"
  ) +
  theme_fce()

ggplotly(p_milk, tooltip = "text") %>%
  layout(hoverlabel = list(bgcolor = "white"))
```

# Executive Summary {data-icon="fa-clipboard-check"}

## Row {data-height=100}

### Reserve Sufficiency {.value-box}
```{r}
valueBox(
  value   = percent(prob_ok, accuracy = 0.1),
  caption = "Probability Current Reserve Is Sufficient",
  icon    = "fa-check-circle",
  color   = fce_green
)
```

### Reserve Buffer {.value-box}
```{r}
valueBox(
  value   = dollar(buffer),
  caption = "Capital Buffer Above 99th Pct Loss",
  icon    = "fa-shield-alt",
  color   = ifelse(buffer > 0, fce_mid, fce_red)
)
```

### CVaR {.value-box}
```{r}
valueBox(
  value   = dollar(cvar_95),
  caption = "Expected Shortfall — Avg Loss in Worst 5%",
  icon    = "fa-exclamation-circle",
  color   = fce_orange
)
```

### Dairy Concentration {.value-box}
```{r}
dairy_share <- mean(sim_results$dairy_loss) / el_mean
valueBox(
  value   = percent(dairy_share, accuracy = 0.1),
  caption = "Dairy Share of Expected Loss (28% of Portfolio)",
  icon    = "fa-tint",
  color   = fce_red
)
```

## Row {data-height=380}

### Board Finding {data-width=420}

<div style="padding: 8px 4px;">

<h3 class="section-header">Monte Carlo Risk Assessment</h3>

<div class="method-box" style="border-color: #2D6A4F;">
<h4>✅ Reserve Adequacy — Confirmed with Caveat</h4>
<p>The current CECL reserve of <strong>$8,100,000</strong> is sufficient in
<strong>`r percent(prob_ok, accuracy = 0.1)`</strong> of all historically
plausible macro scenarios. Under baseline and moderately stressed conditions,
FCE's capital position is sound.</p>
<p><strong>Caveat:</strong> The 99th percentile loss of
<strong>`r dollar(var_99)`</strong> leaves a buffer of only
<strong>`r dollar(buffer)`</strong> — approximately
<strong>`r percent(buffer/cecl_reserve, accuracy = 0.1)`</strong>
of the reserve — before a capital shortfall occurs.</p>
</div>

<div class="method-box" style="border-color: #F4A261;">
<h4>⚠️ Dairy Concentration Risk — Quantified</h4>
<p>Dairy represents <strong>28%</strong> of portfolio exposure but drives
<strong>`r percent(mean(sim_results$dairy_loss)/el_mean, accuracy=0.1)`</strong>
of expected losses — a concentration ratio of
<strong>`r round((mean(sim_results$dairy_loss)/el_mean) / 0.28, 2)`×</strong>.
Current milk price index of <strong>286.9</strong> is favorable,
but a reversal would disproportionately stress reserves.</p>
</div>

<div class="method-box" style="border-color: #C1121F;">
<h4>📋 Board Recommendation</h4>
<p>Monitor the <strong>`r dollar(buffer)`</strong> reserve buffer quarterly.
Consider establishing a trigger at 90th percentile loss
(<strong>`r dollar(quantile(sim_results$total_loss, 0.90))`</strong>)
as an early warning threshold for reserve augmentation review.</p>
</div>

</div>

### Full Results Table {data-width=580}
```{r}
exec_summary <- tibble(
  Category = c(
    rep("Loss Distribution", 6),
    rep("Capital Adequacy", 5),
    rep("Segment Losses", 4),
    rep("Model Inputs", 5)
  ),
  Metric = c(
    "Expected Loss (Mean)",
    "Median Loss (50th Pct)",
    "Elevated Loss (75th Pct)",
    "Value at Risk (95th Pct)",
    "Severe Stress (99th Pct)",
    "Expected Shortfall (CVaR)",
    "Current CECL Reserve",
    "Reserve Covers Expected Loss",
    "Probability Reserve Sufficient",
    "Buffer Above 99th Pct",
    "Recommended Watch Threshold",
    "Dairy Avg Loss",
    "Livestock Avg Loss",
    "Row Crop Avg Loss",
    "Agri-Business Avg Loss",
    "Simulations Run",
    "PD Model AUC",
    "Ag Calibration Factor",
    "Macro Data Period",
    "Training Loans"
  ),
  Value = c(
    dollar(el_mean),
    dollar(el_median),
    dollar(var_75),
    dollar(var_95),
    dollar(var_99),
    dollar(cvar_95),
    dollar(cecl_reserve),
    "YES",
    percent(prob_ok, accuracy = 0.1),
    dollar(buffer),
    dollar(quantile(sim_results$total_loss, 0.90)),
    dollar(mean(sim_results$dairy_loss)),
    dollar(mean(sim_results$livestock_loss)),
    dollar(mean(sim_results$row_crop_loss)),
    dollar(mean(sim_results$agribiz_loss)),
    "10,000",
    "0.7012",
    "0.069 (FRED DRALACBS ÷ LC default rate)",
    "2000–2025",
    "1,327,062 loans"
  )
)

datatable(
  exec_summary,
  rownames = FALSE,
  options  = list(
    dom        = "ft",
    pageLength = 20,
    scrollY    = "310px",
    columnDefs = list(
      list(width = "120px", targets = 0),
      list(width = "220px", targets = 1)
    )
  ),
  class = "compact stripe"
) %>%
  formatStyle(
    "Category",
    fontWeight      = "700",
    fontSize        = "0.78rem",
    color           = fce_dark,
    textTransform   = "uppercase",
    letterSpacing   = "0.05em"
  ) %>%
  formatStyle(
    "Value",
    fontFamily = "DM Mono, monospace",
    fontSize   = "0.83rem"
  ) %>%
  formatStyle(
    "Metric",
    fontSize = "0.83rem"
  )
```